A thermodynamic view of dusty protoplanetary disks

### 2017-05-22 appendix.tex
edits on the plane
fixed tiny typo in appendix
new section 7 on secular GI. rest of paper might need adjusting. mostly doing non-SG, but one sec and one appendix has SG
misc edits from sect 4 to sect 8.3

Here we consider the more general equation of state for the gas with
$P = K \rhog^{\Gamma}$ where $K$ is a prescribed function of position
and $\Gamma$ is a constant. The effective energy equation is then
Eq. \ref{poly_energy}. We also include the self-gravitational
potential $\psi$ of the gas
plus dust mixture, which satisfies 
%\begin{align}
$  \nabla^2 \psi  = 4 \pi G \rho. $
%\end{align} 
The linearized equations for axisymmetric
disturbances are

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
%  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
   \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \hat{\bm{r}} \cdot \delta\bm{F} -  \hat{\bm{r}} \cdot \nabla\dd\psi ,\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \hat{\bm{z}} \cdot \delta\bm{F}  -  \hat{\bm{z}} \cdot \nabla\dd\psi, \\ 
\nabla^2 \delta\psi & = 4\pi G \dd\rho. 
\end{align}  
where the linearizd pressure force $\delta \bm{F}$ is given in
Appendix \ref{lin_press}. These equations do not assume the
radially-local approximation used in the main text. From the
linearized meridional momentum equations, we find

@@

\begin{align}
  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\label{streaming_mass}\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\label{streaming_vx}\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\label{streaming_vz}\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
%{\bf bg pressure gradient neglected in energy eq (is this ok? could
%  fix, but get more complicated dispersion}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. We note that the incompressible
condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. 
\end{align}
%{\bf only the second term contributes to work done? dont think so}

We eliminate the velocity perturbations to obtain
\begin{align}
  \left(\ii \zeta \sigma - \tstop c_s^2 \tepsilon |\bm{k}|^2\right)Q &=
  \ii \left[
  \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
  2\tepsilon\right)\right]W, \\
   \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q, 
\end{align}
which yields the dispersion relation
\begin{align}
  &\frac{\zeta}{c_s^2|\bm{k}|^2}\sigma^5 + \ii \tepsilon \tstop
  \sigma^4 - \left[ \frac{\zeta}{c_s^2|\bm{k}|^2} \left(\kappa^2 - \ii
  k_x F_r\right) + \left(1 - \tepsilon\right)\right]\sigma^3 - \ii
  \tstop \left[\tepsilon \kappa^2 - \ii k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 \notag\\ 
  &+ \left(1 -
  \tepsilon\right)\left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma +
  k_x \tstop F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0.\label{streaming_dispersion}
\end{align}
The equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
was used.

@@

@@

Consider the limit $k_z=0$. Then Eq. \ref{streaming_incompressible}
becomes 
\begin{align}      %streaming_incompressible
\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
\end{align} 
Neglecting the quadratic term leads to stability, $\imag{\sigma}<0$, 
as found by \cite{youdin05a} in their full two-fluid analysis. 
% However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
% \begin{align}
% \omega\left(2 s + \frac{1}{\epsilon\tstop}\right) - \frac{k_xF_r}{\epsilon} & =
%                                                                      0, \label{epi_overs1} \\
% \omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs2}
% \end{align}
% so the growth rate satisfies 
% \begin{align*}
% \left( \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}\right)^2 - s^2
%   - \frac{s}{\epsilon \tstop} = \kappa^2. %\omega = \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}
% \end{align*} 
However, solving Eq. \ref{kz_zero_si} explicitly assuming $\epsilon s \tstop \ll 1$ would yield 
\begin{align}
  |\omega| &\simeq \left|k_xF_r\right| \tstop \sim \kappa, \\
  s &\simeq \epsilon \tstop \left[\left(k_xF_r\tstop\right)^2 - \kappa^2\right],\label{epi_grow}
\end{align}
Accordingly, given $k_xF_r$, growth is possible if 
\begin{align}
\tstop > \frac{\kappa}{k_x\left|F_r\right|}\equiv t_{\mathrm{s}0}.\label{epi_grow_criterion}
\end{align}
The mode is marginally stable when $\tstop =
t_{\mathrm{s}0}$. Alternatively, for fixed $\tstop$ growth is enabled
by a sufficiently large $\left|k_xF_r\right|$. Unlike the streaming instability,
which is strongly surpressed when $\epsilon=1$, these modes can still
grow at equal dust-to-gas ratio.

These growing epicycles with $|\omega|\gtrsim \kappa$ are absent in
full two-fluid models \citep{youdin05a}. The discrepency lies in the
fact that the one-fluid equations, to first order in $\tstop$, are
only valid for low-frequency waves with $|\sigma|\lesssim
\Omega$. Thus only low-frequency modes should be retained from
analyses based on Eqs. \ref{masseq}---\ref{tempeq}.

@@

Fig. \ref{dusty_growth1} show unstable modes found from
Eq. \ref{streaming_dispersion} as a function of $K_z$ at fixed $K_x 
= 1500$, $\epsilon=2$ and $\taus\OmK=0.01$.  
The (spurious) overstable dusty epicycles' growth
rates are weakly depedent on $K_z$ and are well approximated by
that in the $K_z=0$ limit, Eq. \ref{epi_grow}, which becomes 
\begin{align*} 
  s = \epsilon \left(1 - \tepsilon
  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
  -1 \right]
\end{align*}
in the above notation. 
For the case considerd in Fig. \ref{dusty_growth1} we find
$s\simeq0.067\OmK$, as observed. By contrast, the streaming
instability requires $K_z>0$, and dominates when
$K_z\gtrsim 100$.

>>>>

Here we consider the more general equation of state for the gas with
$P = K \rhog^{\Gamma}$ where $K$ is a prescribed function of position
and $\Gamma$ is a constant. The effective energy equation is then
Eq. \ref{poly_energy}. % We also include the self-gravitational
% potential $\psi$ of the gas
% plus dust mixture, which satisfies 
% %\begin{align}
% $  \nabla^2 \psi  = 4 \pi G \rho. $
% %\end{align} 
The linearized equations for axisymmetric
disturbances are

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
%  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
   \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \hat{\bm{r}} \cdot \delta\bm{F} -  \hat{\bm{r}} \cdot \nabla\dd\psi ,\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \hat{\bm{z}} \cdot \delta\bm{F}  -  \hat{\bm{z}} \cdot \nabla\dd\psi, \\ 
\nabla^2 \delta\psi & = 4\pi G \dd\rho. 
\end{align}  
where the linearizd pressure force $\delta \bm{F}$ is given in
Appendix \ref{lin_press}. These equations do not assume the
radially-local approximation used in numerical computations. From the
linearized meridional momentum equations, we find

@@

\begin{align}
  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\label{streaming_mass}\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\label{streaming_vx}\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\label{streaming_vz}\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
%{\bf bg pressure gradient neglected in energy eq (is this ok? could
%  fix, but get more complicated dispersion}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. We note that the incompressible
condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. \label{lin_drag_si}
\end{align}
%{\bf only the second term contributes to work done? dont think so}

We eliminate the velocity perturbations to obtain
\begin{align}
  \left(\ii \zeta \sigma - \tstop c_s^2 \tepsilon |\bm{k}|^2\right)Q &=
  \ii \left[
  \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
  2\tepsilon\right)\right]W, \label{si_QW1}\\
   \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q, 
\end{align}
which yields the dispersion relation
\begin{align}
  &\frac{\zeta}{c_s^2|\bm{k}|^2}\sigma^5 + \ii \tepsilon \tstop
  \sigma^4 - \left[ \frac{\zeta}{c_s^2|\bm{k}|^2} \left(\kappa^2 - \ii
  k_x F_r\right) + \left(1 - \tepsilon\right)\right]\sigma^3 - \ii
  \tstop \left[\tepsilon \kappa^2 - \ii k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 \notag\\ 
  &+ \left(1 -
  \tepsilon\right)\left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma +
  k_x \tstop F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0.\label{streaming_dispersion}
\end{align}
The equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
was used.

@@

\subsection{Approximate solutions in the dust-rich limit}\label{si_dust_rich}
Here we seek analytic solutions for the streaming instability by 
examining limiting cases and with further approximations. We shall fix
$k_z$ and $\fdust>1/2$ (the dust-rich case), and maximize growth rates
over $k_x$. We consider small stopping times, $\tau_\mathrm{s}\to0$,
and assume that the optimum $k_x\to\infty$ but that
$k_x\tau_\mathrm{s}$ remain bounded.

We begin with the incompressible dispersion
relation, Eq. \ref{streaming_incompressible}. We assume a Keplerian
disk and neglect the quartic  
term because it is higher order in $\tau_\mathrm{s}$ (see also
\S\ref{spurious_epi})  In dimensionless form: 
\begin{align}
  \nu^3 + \st\left[\ii\tepsilon + 2 K_x \fg \left(1-\fdust\right)\right]\nu^2
  - \left(\frac{K_z}{K_x}\right)^2\nu - 2\st K_x \fg
  \left(1-2\fdust\right)\left(\frac{K_z}{K_x}\right)^2 = 0,\label{dimless_cubic}
\end{align}
where $\nu = \sigma/\Omega$, $\st = \taus\Omega$, $\fg = 1-\fdust$ (the gas
fraction), and recall $K_{x,z}=k_{x,z}\eta r$. We have used $F_r =
2\eta \fg r \Omega^2$ and considered large $K_x$.

Full analytic expressions for cubic roots are unwieldy. To proceed
in an a tractable manner, let us \emph{assume} at this stage that the
quadratic term can be neglected compare to the last term. This amounts
to assuming a  sufficiently \emph{large} $K_z$ such that :
\begin{align}
\left|\frac{\ii\fdust}{2K_x\fg} + \left(1 -
      \fdust\right)\right| \left|K_x\nu\right|^2 \ll
  K_z^2\left(2\fdust - 1\right) \label{min_kz}
\end{align}
This will be checked later. (For fixed $\fdust$ and $K_x\to\infty$, the
imaginary term can be neglected.)

Eq. \ref{dimless_cubic} now becomes the depressed cubic 
\begin{align}
  \nu^3 - \calP \nu = \calQ, \label{depressed_cubic}
\end{align}
where $\calP$ and $\calQ$ can be read off Eq. \ref{dimless_cubic}. This can be
solved with Vieta's substitution 
\begin{align*}
\nu = \mu + \frac{\calP}{3\mu},
\end{align*}
then Eq. \ref{depressed_cubic} becomes a quadratic for $\mu^3$,
\begin{align}
\mu^6 - \calQ\mu^3 + \frac{\calP^3}{27} = 0. 
\end{align}
The solution is
\begin{align}
  \mu^3 &= \frac{\calQ}{2}\left(1\pm \sqrt{1 -
      \frac{4\calP^3}{27\calQ^2}}\right)\notag\\
  & = \st \fg (1- 2\fdust)\frac{K_z^2}{K_x}\left\{1 \pm \sqrt{1 -
      \frac{K_z^2}{27K_x^2\left[\st K_x \fg
          \left(1-2\fdust\right)\right]^2}}\right\}.  
\end{align}
At this point we assume $K_z$ is sufficiently \emph{small} such that
the term $\propto K_z^2$ inside the square root may be neglected. That
is, 
\begin{align}
  K_z^2 \ll 27 \left(K_x^2\st\right)^2 \left[\fg
          \left(1-2\fdust\right)\right]^2. \label{max_kz}
%27 K_x^2 \left[\st K_x \fg
%          \left(1-2\fdust\right)\right]^2.
\end{align}
This will be checked later. Remembering that we are considering the
dust-rich case so that $1-2\fdust<0$, the solution for $\mu$ is   
\begin{align}
\mu = \left[2\st\fg \left(2\fdust
    -1\right)\frac{K_z^2}{K_x}\right]^{1/3}e^{\ii\pi/3}. \label{mu_explicit} 
\end{align}
We have chosen the root for instability. We can now write down the
solution for the eigenfrequency $\nu$:
\begin{align}
\real{(\nu)} &= \frac{1}{2}\left(|\mu| + \frac{K_z^2}{3
    K_x^2|\mu|}\right)\notag\\
&= \frac{1}{2}\left(\calA \st^{1/3}K_x^{-1/3} + \calB
    \st^{-1/3}K_x^{-5/3}\right),\\
\imag{(\nu)} &= \frac{\sqrt{3}}{2}\left(|\mu| - \frac{K_z^2}{3
    K_x^2|\mu|}\right),\notag\\
&= \frac{\sqrt{3}}{2}\left(\calA \st^{1/3}K_x^{-1/3} - \calB
    \st^{-1/3}K_x^{-5/3}\right),
\end{align}
%Inserting the expression for $|\mu|$ into $\imag{\nu}$, the
%explicit expression for the growth rate may be written as 
%\begin{align}
%  s = \frac{\sqrt{3}}{2}\left(\calA \st^{1/3}K_x^{-1/3} - \calB
%    \st^{-1/3}K_x^{-5/3}\right)\OmK, 
%\end{align}
where
\begin{align}
  &\mathcal{A} = \left[2\fg \left(2\fdust
      -1\right)K_z^2\right]^{1/3} ,\\
  &\mathcal{B} = \frac{K_z^2}{3\calA}. 
\end{align}

Maximizing growth rates over $K_x$ by setting $\p \imag{(\nu)}/\p K_x
= 0$, we find the optium wavenumber is given by
\begin{align}
  K_x = \left(\frac{5\calB}{\calA}\right)^{3/4}\st^{-1/2} =
  \left(\frac{5}{3}\right)^{3/4}\sqrt{\frac{K_z}{2\fg\left(2\fdust-1\right)\st}}. 
\end{align}
The maximum growth rate is then 
\begin{align}
  \mathrm{max}\left(s\right) =
  \frac{2\sqrt{3}\calA^{5/4}}{5^{5/4}\calB^{1/4}}\sqrt{\st}\OmK =
  \frac{2\sqrt{2}\times3^{3/4}}{5^{5/4}}\sqrt{\fg\left(2\fdust-1\right)K_z\st}\OmK.
\end{align}
and the corresponding (dimensional) real frequency satisfies
$\real{(\sigma)} =  \sqrt{3}s/2$.

From the above result we see that as $\taus\to0$, the most unstable
radial wavelength diverges, $K_x\propto \taus^{-1/2} \to \infty$ with
a vanishing growth rate, $\mathrm{max}(s)\propto \sqrt{\taus}\to 0$.

\subsubsection{Finite phase lag as $\taus\to 0$}  
An interesting property of the special solutions described above is
that there is always a phase lag between the Lagrangian pressure and
density perturbations. Since $\real{(\nu)}>0$, the phase lag may be
defined as $\varphi = \arg{\left(\Delta P \Delta \rho^*\right)}$. For
SI we neglect the background density gradients, so $\Delta\rho = 
\delta\rho$. It may be shown that in the limit of large $K_x$ 
that $\Delta P \simeq \ii \delta v_x \p_rP/\sigma$. To obtain $\dd v_x$ we use Eq. \ref{streaming_vx} 
assuming low frequency modes, $|\sigma|\ll \kappa^2$, valid for
$\taus\to 0$ as shown above. Now, the expression for $\dd v_x$ involves
the Eulerian pressure perturbation, $\delta P$. For this we use
Eq. \ref{si_QW1} and assume incompressibility.

With these simplifications (in addition to those already made in the
previous section), the phase lag $\varphi$ is given via 
\begin{align}
  \tan{\varphi} = \frac{\imag{\left(\Delta
        P\Delta\rho^*\right)}}{\real{\left(\Delta
        P\Delta\rho^*\right)}}\simeq\frac{K_x\imag{(\nu)}}{2\fg^2\st
      K_x^2 + K_x \real{(\nu)}}. 
\end{align}
From here it is clear for the solutions derived above, where
$K_x\propto \st^{-1/2}$ and $\nu \propto \st^{1/2}$, that $\varphi$ is
a constant. Explicitly inserting the solutions give 
\begin{align}
\tan{\varphi} = \frac{2\times 3^{3/2}\left(2\fdust-1\right)}{16 -
  7\fdust}. 
\end{align}
For the case shown in Fig. \ref{si_compare_fig} with $\fdust=0.75$, we
have $\varphi\simeq 26\degr$, comparable to the $\sim 30\degr$
obtained from numerical solutions.

\subsubsection{Validty of the approximations}
The solution above require $K_z$ be sufficiently large for
Eq. \ref{min_kz} but also sufficiently small for
Eq. \ref{max_kz}. Inserting the above solutions, we find the
requirement from Eq. \ref{min_kz} becomes 
\begin{align}
  \fdust \gg \frac{12}{17}\simeq 0.7.
\end{align}
However, since $\fdust \leq 1$, this requirement can only be marginally
satisfied. We find that this mostly give errors in the real
frequency (see Fig. \ref{si_compare_fig}), while growth rates are
still captured. On the other hand, Eq. \ref{max_kz} becomes the trivial 
inequality \begin{align}
  1\ll \frac{27}{4}\left(\frac{5}{3}\right)^3\simeq 30,
\end{align}
so Eq. \ref{max_kz} is satisified, which justifies the solution for
$\mu$ in Eq. \ref{mu_explicit}.

@@

Consider the limit $k_z=0$. Then Eq. \ref{streaming_incompressible}
becomes 
\begin{align}      %streaming_incompressible
%\ii \sigma^2 - \frac{\sigma}{\epsilon \tstop}  - \left(\ii \kappa^2 +
%  \frac{k_xF_r}{\epsilon}\right) = 0. \label{kz_zero_si}
\ii\nu^2 - \frac{\nu}{\fdust\st} - \left[\ii + 
  2K_x\frac{\fg}{\fdust}\left(1-\fdust\right)\right]=0\label{kz_zero_si} 
\end{align} 
in dimensionless form. Neglecting the quadratic term leads to
stability, $\imag{(\nu)}<0$, as found by \cite{youdin05a} in their full
two-fluid analysis.  
% However, the real and imaginary parts of Eq. \ref{kz_zero_si}  are
% \begin{align}
% \omega\left(2 s + \frac{1}{\epsilon\tstop}\right) - \frac{k_xF_r}{\epsilon} & =
%                                                                      0, \label{epi_overs1} \\
% \omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs2}
% \end{align}
% so the growth rate satisfies 
% \begin{align*}
% \left( \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}\right)^2 - s^2
%   - \frac{s}{\epsilon \tstop} = \kappa^2. %\omega = \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}
% \end{align*} 
However, solving Eq. \ref{kz_zero_si} explicitly assuming $\left|\fdust\taus
\imag{(\nu)}\right| \ll 1$ would yield  
\begin{align}
%  |\omega| &\simeq \left|k_xF_r\right| \tstop \sim \kappa, \\
%  s &\simeq \epsilon \tstop \left[\left(k_xF_r\tstop\right)^2 - \kappa^2\right],\label{epi_grow}
\real{(\nu)}&\simeq 2K_x\left(1-\fdust\right)^2\st \sim O(1)\\
\imag{(\nu)}&\simeq \fdust \st \left[4K_x^2\left(1-\fdust\right)^4\st^2 -
1\right].\label{epi_grow}
\end{align}
Accordingly, growth is possible if 
\begin{align}
%\tstop > \frac{\kappa}{k_x\left|F_r\right|}\equiv t_{\mathrm{s}0}.\\%\label{epi_grow_criterion}
\st > \frac{1}{2K_x\left(1-\fdust\right)^2}\label{epi_grow_criterion}
\end{align}
%The mode is marginally stable when $\tstop =
%t_{\mathrm{s}0}$. 
Alternatively, for fixed $\taus$ growth is enabled
by a sufficiently large $K_x$. Unlike the streaming instability,
which is strongly surpressed when $\fdust=1/2$, these modes can still
grow at equal dust-to-gas ratio.

These growing epicycles with $|\omega|\gtrsim \kappa$ are absent in
full two-fluid models \citep{youdin05a}. The discrepency lies in the
fact that the one-fluid equations, to first order in $\taus$, are
only valid for low-frequency waves with $|\sigma|\lesssim 
\Omega$. Thus only low-frequency modes should be retained from
analyses based on Eqs. \ref{masseq}---\ref{tempeq}.

@@

Fig. \ref{dusty_growth1} show unstable modes found from
Eq. \ref{streaming_dispersion} as a function of $K_z$ at fixed $K_x 
= 1500$, $\epsilon=2$ and $\taus\OmK=0.01$.  
The (spurious) overstable dusty epicycles' growth
rates are weakly depedent on $K_z$ and are well approximated by
that in the $K_z=0$ limit, Eq. \ref{epi_grow}. 
%, which becomes 
%\begin{align*} 
%  s = \epsilon \left(1 - \tepsilon
%  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
%  -1 \right]
%\end{align*}
%in the above notation. 
For the case considerd in Fig. \ref{dusty_growth1} we find
$s\simeq0.067\OmK$, as observed. By contrast, the streaming
instability requires $K_z>0$, and dominates when
$K_z\gtrsim 100$.

### 2017-05-22 linear.tex
home/work sync

In Table \ref{si_compare} we also calculate the phase difference
between the Lagragian pressure perturbation and density perturbations
as    
\begin{align*} 
\varphi \equiv -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right)
        = \sgn(\omega)\arg\left(\Delta\rhog^*\Delta\rhod\right),\notag
\end{align*}
%where we have applied the isothermal equation of state for the second equality.
then $\varphi > 0 $ indicates gas density lagging behind dust density,
which is true for all the cases. In particular, `linB' indicates that
growth rates vanish as $\varphi \to 0$.  Thus the streaming
instability is indeed associated with a phase lag between gas and dust
density evolutions.

In a future work we will perform a detailed comparison between the
simplified one-fluid and full two-fluid frameworks in calculating SI.

\begin{figure}
\includegraphics[width=linewidth]{figures/si_2f_1f_compare}
\caption{{\bf in case the paper isn't long enough} Comparison of the linear
streaming instability obtained from a full two-fluid analysis and the current 
one-fluid framework simplified by the terminal velocity approximation. The vertical
wavenumber is fixed and growth rates are maximized over $K_x$. 
\label{si_compare}}
\end{figure}
%"generalized stokes number" would include other problem parameters,
%e.g. wavenumber

>>>>

In Table \ref{si_compare} we also calculate the phase difference
between the Lagragian pressure perturbation and density perturbations
as    
\begin{align*} 
\varphi \equiv -\sgn(\omega)\arg\left(\Delta P\Delta\rho^*\right)
        = \sgn(\omega)\arg\left(\Delta\rhog^*\Delta\rhod\right),\notag
\end{align*}
%where we have applied the isothermal equation of state for the second equality.
then $\varphi > 0 $ indicates gas density lagging behind dust density,
which is true for all the cases. %In particular, `linB' indicates that
%growth rates vanish as $\varphi \to 0$.  
Thus the streaming instability is indeed associated with a phase lag
between gas and dust density evolutions.

In Fig. \ref{si_compare_fig} we compare the most unstable SI 
mode as a function of $\taus$ at fixed $K_z=30$ and $\epsilon =3$. The
solid line is obtained from a full, two-fluid linear analysis 
\citep[similar to][]{youdin05a, kowalik13}. The one-fluid results are
obtained by solving the dispersion relation
Eq. \ref{streaming_dispersion}, both numerically (green asterisks) and
analytically with additional approximations and considerations
(organge diamonds). The analytic model is described in Appendix 
\ref{si_dust_rich}.

Both results based on the one-fluid framework compares well with the 
full, two-fluid analysis, only breaking down at a relatively large
$\taus\OmK\gtrsim 0.1$. It is interesting to see for large $\taus$ the 
two-fluid phase lag drops, along with the growth rate. Notcie also the
phase lag remains finite as $\taus\to0$ because the optimum $K_x$
diverges (this is explained in Appendix \ref{si_dust_rich}).

In a future work we will perform a more detailed parameter survey to
compare the simplified one-fluid and full two-fluid frameworks in
calculating SI.

\begin{figure}
\includegraphics[width=\linewidth]{figures/si_2f_1f_compare}
\caption{{\bf in case the paper isn't long enough} Comparison of the linear
streaming instability between a full two-fluid analysis (solid
line), the current one-fluid framework simplified by the terminal
velocity approximation (green asterisks) and an analytic solution to
the one-fluid dispersion relation in the dust-rich limit (organge
diamonds, see also Appendix \ref{si_dust_rich}). The vertical 
wavenumber is fixed and growth rates are maximized over $K_x$. 
\label{si_compare_fig}}
\end{figure}
%"generalized stokes number" would include other problem parameters,
%e.g. wavenumber
